\vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}

### 2021-03-22 final 3.Rmd
small changes

---
title: \vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}
author:
- Cole Juracek (Programmer)
- Lauren Palazzo (Writer)
- Lingyu Zhou (Checker + Coordinator)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  pdf_document: default
fontsize: 11pt
geometry: margin=2cm
---

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(pacman)
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr)
# devtools::install_github("goodekat/redres")
```

@@

Among the data available to us are location data pertaining to the city, state, and United States region of the sale. These variables present a natural grouping structure for the price data, which suggest that we employ a hierarchical model in our analysis. The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis. 
# Exploratory data analysis and data cleaning

@@

```{r, echo = FALSE}
# recode sources 
# remove unused levels
df$source <- droplevels(df$source)
source_df <- data.frame(table(df$source))
# levels(df$source)
levels(df$source)[1] <- "Blank"
source_df <- data.frame(table(df$source))
df$source <- recode_factor(df$source, "Blank" = "Blank",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet Pharmacy",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL
source_df_2 <- data.frame(table(df$source))
# drop unknowns
df <- df %>% filter(df$source != "Unknown")
df$source <- droplevels(df$source)
knitr::kable(table(df$source), col.names = c("Level", "Freq"))
```

@@

```{r}
#table(df$bulk_purchase, useNA = "always")
# looks fine
```

@@

```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)
reason_df <- data.frame(table(df$Primary_Reason))
# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Blank"
reason_df <- data.frame(table(df$Primary_Reason))
df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
                                   "0 Reporter did not answer the question" = "Left unanswered",
                                   "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                                   "5 To resell" = "Resell",
                                   "8 Prefer not to answer" = "Refuse to answer",
                           .default = "Other or unknown")
reason_df_2 <- data.frame(table(df$Primary_Reason))
#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

@@

Model selection is omitted for brevity (see code appendix for full demonstration). Essentially we compared nested models with fewer and fewer variables, provided the results of the Anova test suggested the simpler model was preferred. Ultimately, this led to us dropping `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source

@@

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
# resqq <- plot_resqq(mod1)
# ranefci <- plot_ranef(mod1)
# grid.arrange(resqq, ranefci, ncol = 2)
```

Model diagnostics discussions here.

## Results
#### Fixed Effects
```{r, message=FALSE}
#knitr::kable(fixef(mod1))
#confint(mod1)[3:11,]
# extract coefficients
coefs <- data.frame(coef(summary(mod_final)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
```

*dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone increases by a multiplicative effect of exp(-0.0151868) = 0.9849279, about a 1.51% reduction.

*bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1691121) = 0.8444142, about a 15.56% reduction.

*source* (“Source: Unknown” is the reference, other levels ommitted for brevity): Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1339468) = 1.143332, about a 14.33% increase than the price per mg of oxymorphone for people with unknown drug source.

*Primary Reason* ("Blank" is the reference, other levels omitted for brevity): Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1034974) = 1.109043, about a 10.9% increase than the price per mg of oxymorphone for people with Blank primary reason.

#### Random Effects
```{r}
#VarCorr(mod_final)
Variance <- c(0.02292, 0.50060)
Std.Dev <- c(0.15140 , 0.70753)
Groups <- c("state (Intercept)", "Residual")
REvar <- cbind(Groups, Variance, Std.Dev)
knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
```
The within-state variance $\sigma^2$ is 0.5006, which is much larger than the across-state variance $\tau^2$, 0.02292.

@@

```{r, fig.align='center', fig.height=8}
dotplot(ranef(mod_final, condVar=TRUE))$state
```
From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero. The random intercept for Arizona is the lowest and this intercept is also significant.

@@

### Date variable  
```{r}
hist(df$date_num)
knitr::kable(table(df$date_quarter))
```

### Dose strength variable 
```{r}
hist(df$dose_per_mg)
```

@@

### City visualization for FL
```{r}
ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")
```

@@

group_int_list <- NULL 
group_slope_list <- NULL
for(statej in unique(df$state)) {
    m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
    group_int_list <- c(group_int_list, coef(m)[1])
    group_slope_list <- c(group_slope_list, coef(m)[2])
}
hist(group_int_list)
hist(group_slope_list)

>>>>

---
title: "\vspace{-1.5cm} \textbf{STA610 Case Study 1 Team 4 Report}"
author:
- Cole Juracek (Programmer)
- Lauren Palazzo (Writer)
- Lingyu Zhou (Checker + Coordinator)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
fontsize: 11pt
geometry: margin=2cm
---

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(pacman)
library(merTools)
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr,
               influence.ME, knitr)
# devtools::install_github("goodekat/redres")
```

@@

Among the data available to us are location data pertaining to the city, state, and United States region of the sale. These variables present a natural grouping structure for the price data, which suggest that we employ a hierarchical model in our analysis. The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.

# Exploratory data analysis and data cleaning

@@

```{r, echo = FALSE}
# recode sources 
# remove unused levels
df$source <- droplevels(df$source)
source_df <- data.frame(table(df$source))
# levels(df$source)
levels(df$source)[1] <- "Blank"
source_df <- data.frame(table(df$source))
df$source <- recode_factor(df$source, "Blank" = "Blank",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet Pharmacy",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL
source_df_2 <- data.frame(table(df$source))
# drop unknowns
df <- df %>% filter(df$source != "Unknown")
df$source <- droplevels(df$source)
knitr::kable(table(df$source), col.names = c("Level", "Freq"), caption = "Frequency Table of Variable source")
```

@@

@@

```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)
reason_df <- data.frame(table(df$Primary_Reason))
# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Blank"
reason_df <- data.frame(table(df$Primary_Reason))
df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
                                   "0 Reporter did not answer the question" = "Left unanswered",
                                   "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                                   "5 To resell" = "Resell",
                                   "8 Prefer not to answer" = "Refuse to answer",
                           .default = "Other or unknown")
reason_df_2 <- data.frame(table(df$Primary_Reason))
#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"), caption = "Frequency Table of Variable Primary_Reason")
```

@@

Model selection is omitted for brevity (see code appendix for full demonstration). Essentially we compared nested models with fewer and fewer variables, provided the results of the Anova test suggested the simpler model was preferred. Ultimately, this led to us dropping `date_num` and `date_quarter`, while keeping: dose_per_mg, bulk_purchase, primary_reason, source.

@@

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
resqq <- plot_resqq(mod_final)
ranefci <- plot_ranef(mod_final)
grid.arrange(resqq, ranefci, ncol = 2)
```

From the normal quantile plot of the raw conditional residuals, although there are many residuals out of bounds, these residuals are not too far away from the bounds. Thus, we can still accept the raw conditional residuals are normally distributed. 
From the normal quantile plot on random effect state, there are only two outliers. Thus, we can accept random effect of state is normally distributed.

## Results

#### Fixed Effects

```{r, message=FALSE}
coefs <- data.frame(coef(summary(mod_final)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
```

* *dose_per_mg*: Controlling for other variables, for every one milligram increase of dosage, the price per mg of oxymorphone reduces by a multiplicative effect of exp(-0.0152224) = 0.9848929, about a 1.51% reduction.

* *bulk_purchase* ("not bulk purchase" is the reference): Controlling for other variables, bulk purchase reduces the price per mg of oxymorphone by a multiplicative effect of exp(-0.1716889) = 0.8422412, about a 15.76% reduction than the price per mg of oxymorphone for people without bulk purchase.

* *source* (“Source: Unknown” is the reference, other levels ommitted for brevity):
    + Controlling for other variables, purchasing oxymorphone with Internet source tends to increase the price per mg by a multiplicative effect of exp(0.1361942) = 1.145904, about a 14.59% increase than the price per mg of oxymorphone for people with unknown drug source.

* *Primary Reason* ("Blank" is the reference, other levels omitted for brevity):
    + Controlling for other variables, purchasing oxymorphone with primary reason Self-treat tends to increase the price per mg by a multiplicative effect of exp(0.1042256) = 1.109851, about a 10.99% increase than the price per mg of oxymorphone for people with Blank primary reason.

#### Random Effects

```{r, include=FALSE}
VarCorr(mod_final)
```

```{r}
Variance <- c(0.02292, 0.50060)
Std.Dev <- c(0.15140 , 0.70753)
Groups <- c("state (Intercept)", "Residual")
REvar <- cbind(Groups, Variance, Std.Dev)
knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))
```
The within-state variance $\sigma^2$ is 0.5006, which is much larger than the across-state variance $\tau^2$, 0.02292.

@@

```{r, fig.align='center', fig.height=8}
plotREsim(REsim(mod_final,n.sims=100),stat='median',sd=TRUE, labs = TRUE)
```
From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero. The random intercept for Arizona is the lowest and this intercept is also significant.

## Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

@@

### Date variable  
```{r}
hist(df$date_num, xlab = "data_num", main = "Histogram of date_num")
knitr::kable(table(df$date_quarter), caption = "Frequency Table of Quarter in date_num")
```

### Dose strength variable 
```{r}
hist(df$dose_per_mg, xlab = "does_per_mg", main = "Histogram of dose_per_mg")
```

@@

### City visualization for FL
```{r, fig.height=10}
ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
  geom_boxplot() +
  labs(title="log PPM by city (FL)", x="Log PPM",y="city")
```

@@

group_int_list <- NULL 
group_slope_list <- NULL
for(statej in unique(df$state)) {
    m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
    group_int_list <- c(group_int_list, coef(m)[1])
    group_slope_list <- c(group_slope_list, coef(m)[2])
}
par(mfrow=c(1,1))
hist(group_int_list)
hist(group_slope_list)
